---
title: "Assignment 5"
author: "Jenna Morris"
date: "12 February 2021"
output: 
  html_document:
    theme: simplex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center", out.width = '70%')
```


## Exploring the the Siscowet Lake Trout dataset from [{FSAdata}](https://github.com/droglenc/FSAdata)


### Load libraries

```{r load_libraries, message=FALSE}
library(here)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
```

### Set directories & load data
```{r load_data, message=FALSE}
data_dir = here::here("data")
fish = read_csv(file.path(data_dir, "siscowet.csv"))
```

### Look at the data
```{r print_data, message = FALSE}
print(fish)
```


### Let's explore the relationship between length `len` and weight `wgt`

**Looking at histograms of each variable:**
```{r histograms, message = FALSE}
hist(fish$len)
hist(fish$wgt)
```

Looks like there's an outlier at the upper range for `wgt`.

### Let's take a look at an exploratory scatterplot:
```{r scatterplot, message = FALSE}
fish %>%
  ggplot(aes(x = len, y = wgt)) +  
  geom_point()
```

Looks like there's a non-linear positive relationship between `len` and `wgt`. We can also see the `wgt` outlier.

**Let's remove the outlier and any NA values:**
```{r remove_outlier, message = FALSE}
fish_sub = fish %>%
  dplyr::filter(wgt < 5000)
```

**Let's compare the relationship across sex:**
```{r sex, message = FALSE}
fish_sub %>%
  ggplot(aes(x = len, y = wgt, color = sex)) +  # sex as color
  geom_point() 
```
The relationship appears to be similar for both male and female trout.


**Let's compare the relationship across location:**
```{r location, message = FALSE}
fish_sub %>%
  ggplot(aes(x = len, y = wgt, color = locID)) +  # location as color
  geom_point()

# location as facet
fish_sub %>%
  ggplot(aes(x = len, y = wgt, color = locID)) +
  geom_point() +
  facet_wrap(~locID)
```
The relationship also appears to be similar across all four locations.


### Let's create an expository figure 

I want to show the relationship between length `len` and weight `wgt` and highlight a specific location of interest.

**Let's clean up the exploratory plot:**
```{r clean_plot, message = FALSE}
# creating a base plot with our variables defined
len_wgt = ggplot(fish_sub, aes(x = len, y = wgt, col = locID))


len_wgt + 
  # using a jittered scatterplot to show points clearly
  geom_jitter(shape = 16, alpha = 0.5, size = 2) +
  facet_wrap(~locID) +
  theme(
    # creating a clean background with light outline and no legend
    panel.background = element_blank(),
    panel.border = element_rect(colour = "gray80", fill = NA),
    legend.position = "none"
  ) +
  # adjusting scales so both start at 0
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     limits = c(0, NA)) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 800)) +
  # adding axis labels
  xlab("Total length (mm)") +
  ylab("Weight (g)") +
  #changing the colors to something more aesthetically pleasing and colorblind friendly
  scale_color_manual(values = c("#274d52", "#c7a2a6", "#818b70", "#604e3c"))
```

**Let's get rid of the facet, and use color instead to highlight the relationship at the Deer Park location:**
```{r location_specific, message = FALSE}
#subset fish from Deer Park
dp = fish_sub %>%
  dplyr::filter(locID == "Deer Park")

# subset fish from all other locations
other = fish_sub %>%
  dplyr::filter(locID != "Deer Park")

ggplot() +
  geom_jitter(data = other, aes(x = len, y = wgt), color = "#878787", alpha = 0.5, shape = 16, size = 2) +
  geom_jitter(data = dp, aes(x = len, y = wgt), color = "#3c6fde", alpha = 0.3, shape = 16, size = 2) +
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(colour = "gray80", fill = NA),
    legend.position = "none"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     limits = c(0, NA)) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 800)) +
  xlab("Total length (mm)") +
  ylab("Weight (g)") 
```

**Let's add a title and location key:**
```{r add_title}
ggplot() +
  geom_jitter(data = other, aes(x = len, y = wgt), color = "#878787", alpha = 0.5, shape = 16, size = 2) +
  geom_jitter(data = dp, aes(x = len, y = wgt), color = "#3c6fde", alpha = 0.3, shape = 16, size = 2) +
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(colour = "gray80", fill = NA),
    legend.position = "none"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     limits = c(0, NA)) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 800)) +
  xlab("Total length (mm)") +
  ylab("Weight (g)") +
  labs(color="Location",
     title = "Trout weight increases with total length") +
  annotate(geom = "text", x = 350, y = 2000, 
                label = "Deer Park", 
                color = "#3c6fde") +
  annotate(geom = "text", x = 700, y = 1300,
                label = "Blind Sucker",
                color = "#878787") +
  annotate(geom = "text", x = 700, y = 1100,
                label = "Grand Marais",
                color = "#878787")+
  annotate(geom = "text", x = 725, y = 900,
                label = "Little Lake Harbour",
                color = "#878787")
```


### Final plot
```{r final_plot, echo = FALSE, out.width = '90%'}
final = ggplot() +
  geom_jitter(data = other, aes(x = len, y = wgt), color = "#878787", alpha = 0.5, shape = 16, size = 2) +
  geom_jitter(data = dp, aes(x = len, y = wgt), color = "#3c6fde", alpha = 0.3, shape = 16, size = 2) +
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(colour = "gray80", fill = NA),
    legend.position = "none",
    plot.subtitle = element_text(color = "gray25", face = "italic")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     limits = c(0, NA)) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 800)) +
  xlab("Total length (mm)") +
  ylab("Weight (g)") +
labs(color="Location",
     title = "Trout weight increases with total length",
     subtitle = "Siscowet Lake Trout captured at Deer Park share similar physiological traits\nwith those caught at other Michigan locations on Lake Superior") +
  annotate(geom = "text", x = 350, y = 2000, 
                label = "Deer Park", 
                color = "#3c6fde") +
  annotate(geom = "text", x = 700, y = 1300,
                label = "Blind Sucker",
                color = "#878787") +
  annotate(geom = "text", x = 700, y = 1100,
                label = "Grand Marais",
                color = "#878787")+
  annotate(geom = "text", x = 725, y = 900,
                label = "Little Lake Harbour",
                color = "#878787")
final
```
